---
title: "Preparing MARINe raw data for power analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, error=FALSE)
```

## Load packages
```{r}
library(tidyverse)
library(janitor)
library(here)
library(vegan)
```

# SWATH DATA
## Quick data tidy
```{r}
swath <- read_csv(here("data","amelia_ritger_cbs_ltm_data_20210601-swath.csv")) %>% 
  clean_names() %>% 
    rename("species" = "final_classification",
         "site"="intertidal_sitename") %>%
  select("site","site_code","survey_rep","year","month","transect","location","species","count")

sort(unique(swath$species))
```

## Check out summary data
```{r}
swath_clean <- swath %>% 
  filter(species!="no swath species found") %>% 
  group_by(site, year, species) %>% 
  summarize(tot_sp=sum(count))
```

## Look into some high numbers
```{r}
swath_lol <- swath %>% 
  filter(species!="no swath species found",
         site=="Lompoc Landing",
         year=="2014")

#Confirm with MARINe summary data for swath counts
marine <- read_csv(here("data","amelia_ritger_cbs_summary_2020_1005-swath.csv")) %>% 
  clean_names() %>% 
  filter(intertidal_sitename=="Lompoc Landing",
         year=="2014")
```

Yep, my quick wrangling matches up with the MARINe summarized data 

## Quick visualization 
```{r}
ggplot(swath_clean, aes(fill=species, y=tot_sp, x=year)) + 
  geom_bar(position="fill", stat="identity") +
  facet_wrap(~site)

#Look at it each year
ggplot(swath_clean, aes(fill=species, y=tot_sp, x=site)) + 
  geom_bar(position="fill", stat="identity") +
  facet_wrap(~year)
#+facet_wrap(~site)
```

I want to exclude Cabrillo (it's SO different from the other sites) and compare only years where all three other sites were measured - so 2016, 2018, 2019, and 2020

```{r}
swath_three <- swath_clean %>% 
  filter(site!="Cabrillo III",
         year %in% c(2016, 2018, 2019, 2020))

ggplot(swath_three, aes(fill=species, y=tot_sp, x=site)) + 
  geom_bar(position="fill", stat="identity")

```

### Damn, check out that (lack of diversity)!

# QUADRAT DATA
## Quick data tidy
```{r}
quad <- read_csv(here("data","amelia_ritger_cbs_ltm_data_20210601-quad.csv")) %>% 
  clean_names() %>% 
    rename("species" = "final_classification",
         "site"="intertidal_sitename") %>%
  select("site","survey_rep","year","month","transect","location","zone_code","species","count")
```

## Check out summary data
```{r}
sort(unique(quad$species))

quad_clean <- quad %>% 
  filter(species %in% c("acanthinucella spp","cancer antennarius","cancer spp", "kelletia kelletii", "leptasterias spp","nucella canaliculata","ucella emarginata; nucella ostrina","pachygrapsus crassipes","hemigrapsus nudus","pisaster ochraceus","tegula aureotincta","tegula brunnea","tegula eiseni","tegula funebralis","tegula gallina")) %>% 
  group_by(site, year, species) %>% 
  summarize(tot_sp=sum(count))
```
# In the future (when I have time) - check species list for predators I may have missed

## check it with MARINe summarized data
```{r}
marine <- read_csv(here("data", "amelia_ritger_cbs_summary_2020_1005-quad.csv")) %>% 
  clean_names() %>% 
  filter(intertidal_sitename=="Alegria",
         year=="2001")
```

Yep, looks good

## Plot it up 
```{r}
ggplot(quad_clean, aes(fill=species, y=tot_sp, x=site)) + 
  geom_bar(position="fill", stat="identity") + #facet_wrap(~year)
  scale_x_discrete(limits = c("Bodega", "Lompoc Landing", "Alegria", "Cabrillo III"))

ggsave("figures/composition_site.png")

```

## Run an NMDS on the data to compare community composition across sites
```{r}
#example community-by-species matrix, nmds
set.seed(2)
community_matrix=matrix(
   sample(1:100,300,replace=T),nrow=10,
   dimnames=list(paste("community",1:10,sep=""),paste("sp",1:30,sep="")))

example_NMDS=metaMDS(community_matrix, # Our community-by-species matrix
                     k=2) # The number of reduced dimensions

#do it with my data
quad_matrix_df <- quad %>%
  filter(species %in% c("acanthinucella spp","cancer antennarius","cancer spp", "kelletia kelletii", "leptasterias spp","nucella canaliculata","nucella emarginata; nucella ostrina","pachygrapsus crassipes","hemigrapsus nudus","pisaster ochraceus","tegula aureotincta","tegula brunnea","tegula eiseni","tegula funebralis","tegula gallina")) %>%
  group_by(site, species) %>% 
  summarize(tot_sp=sum(count)) %>% 
  pivot_wider(names_from=species, values_from=tot_sp) %>% 
  replace(is.na(.), 0)

#convert df to matrix for NMDS
quad_matrix <- data.matrix(quad_matrix_df)
rownames(quad_matrix) <- c("Alegria", "Bodega", "Cabrillo III", "Lompoc Landing") #rename matrix row names to sites
quad_matrix <- quad_matrix[,-1] #remove first column

#run the NMDS
quad_nmds <- metaMDS(quad_matrix, k=2) #k is the number of reduced dimensions
quad_nmds

stressplot(quad_nmds)
plot(quad_nmds, type="t")
```

# Run power analysis on only species found at all sites across sites

## Nucella
```{r}
quad_nucella <- quad %>%
  filter(species %in% c("nucella emarginata; nucella ostrina"),
         survey_rep<ifelse(site=="Cabrillo III", 5, 2)) %>% #Cabrillo only counted nucella on the 4th replicate of surveys so make ifelse
  group_by(site) %>% 
  summarize(tot=n(),
            avg=mean(count),
            var=var(count))

nucella_avg <- quad_nucella %>% 
  pull(avg)
nucella_var <- quad_nucella %>% 
  pull(var) %>% 
  mean()
nucella_n <- quad_nucella %>% 
  pull(tot) %>% 
  mean() %>% 
  ceiling() #round up to whole number for simulation
```

Use simulation to run power analysis for comparing nucella across sites
```{r}
n.sim <- 10000         ## number of simulations
mu    <- nucella_avg  ## site means
sigma2<- nucella_var  ## error variance (averaged across all sites) assumption: 7.5
n     <- nucella_n    ## number of observations per site (mean across all years: 31)

n     <- 17           # coerce new number for replicate size
g     <- length(mu) 
group <- factor(rep(LETTERS[1:g], each = n))

results <- numeric(n.sim) ## vector to store results in

for(i in 1:n.sim){
  ## Simulate new response, build data set
  y <- rnorm(n * g, mean = rep(mu, each = n), sd = sqrt(sigma2))
  data <- data.frame(y = y, group = group)
 
  ## Fit one-way ANOVA model
  fit  <- aov(y ~ group, data = data)
  
  ## Extract result of global F-test
  results[i] <- summary(fit)[[1]][1, "Pr(>F)"] < 0.05 ## 1 = reject
}

mean(results) ## proportion of simulation runs that rejected H_0
```
n = 17, 82%

## Pachygraspus
```{r}
quad_pachy <- quad %>%
  filter(species %in% c("pachygrapsus crassipes"),
         survey_rep==ifelse(site=="Cabrillo III", 4, 
                            ifelse(site=="Bodega", 7,
                                   ifelse(site=="Alegria", 6,1)))) %>%
  group_by(site) %>% 
  summarize(tot=n(),
            avg=mean(count),
            var=var(count))

pachy_avg <- quad_pachy %>% 
  pull(avg)
pachy_var <- quad_pachy %>% 
  pull(var) %>% 
  mean()
pachy_n <- quad_pachy %>% 
  pull(tot) %>% 
  mean() %>% 
  ceiling() #round up to whole number for simulation
```

Use simulation to run power analysis for comparing pachygraspus across sites
```{r}
n.sim <- 10000         ## number of simulations
mu    <- pachy_avg  ## site means
sigma2<- pachy_var  ## error variance (averaged across all sites) assumption: 7.5
n     <- pachy_n    ## number of observations per site (mean across all years: 31)

n     <- 25           # coerce new number for replicate size
g     <- length(mu) 
group <- factor(rep(LETTERS[1:g], each = n))

results <- numeric(n.sim) ## vector to store results in

for(i in 1:n.sim){
  ## Simulate new response, build data set
  y <- rnorm(n * g, mean = rep(mu, each = n), sd = sqrt(sigma2))
  data <- data.frame(y = y, group = group)
 
  ## Fit one-way ANOVA model
  fit  <- aov(y ~ group, data = data)
  
  ## Extract result of global F-test
  results[i] <- summary(fit)[[1]][1, "Pr(>F)"] < 0.05 ## 1 = reject
}

mean(results) ## proportion of simulation runs that rejected H_0
```
n = 25, 81%

## Tegula
```{r}
quad_tegula <- quad %>%
  filter(species %in% c("tegula funebralis"),
         survey_rep==ifelse(site=="Cabrillo III", 4, 
                            ifelse(site=="Bodega", 7,
                                   ifelse(site=="Alegria", 4, 1)))) %>%
  group_by(site) %>% 
  summarize(tot=n(),
            avg=mean(count),
            var=var(count))

tegula_avg <- quad_tegula %>% 
  pull(avg)
tegula_var <- quad_tegula %>% 
  pull(var) %>% 
  mean()
tegula_n <- quad_tegula %>% 
  pull(tot) %>% 
  mean() %>% 
  ceiling() #round up to whole number for simulation
```

Use simulation to run power analysis for comparing tegula across sites
```{r}
n.sim <- 10000         ## number of simulations
mu    <- tegula_avg  ## site means
sigma2<- tegula_var  ## error variance (averaged across all sites) assumption: 7.5
n     <- tegula_n    ## number of observations per site (mean across all years: 31)

n     <- 26           # (if I am running two sets of surveys in a year)
g     <- length(mu) 
group <- factor(rep(LETTERS[1:g], each = n))

results <- numeric(n.sim) ## vector to store results in

for(i in 1:n.sim){
  ## Simulate new response, build data set
  y <- rnorm(n * g, mean = rep(mu, each = n), sd = sqrt(sigma2))
  data <- data.frame(y = y, group = group)
 
  ## Fit one-way ANOVA model
  fit  <- aov(y ~ group, data = data)
  
  ## Extract result of global F-test
  results[i] <- summary(fit)[[1]][1, "Pr(>F)"] < 0.05 ## 1 = reject
}

mean(results) ## proportion of simulation runs that rejected H_0
```
n = 26, 80%

Traditional ANOVA power test (let's use Tegula as an example)
```{r}
library(pwr)
pwr.anova.test(k = 4, f = 0.25, sig.level = 0.01 , power = 0.8) # 25% effect size
```
Damn, need an n = 63 for 80% power

So... at least using my janky power analyses methods, I'll need to do all 11 quadrat replicates to measure significant differences *between* sites

Now let's look at replication needed within a site to capture sufficient variation

# Run power analysis on Tegula (since it has the most intra-site variation) for each site

## Bodega
```{r}
quad_intra_b <- quad %>%
  filter(species %in% c("tegula funebralis"),
         site == "Bodega") %>% 
  group_by(year) %>% 
  summarize(tot=n(),
            avg=mean(count),
            var=var(count))

tegula_avg <- quad_intra_b %>% 
  pull(avg)
tegula_var <- quad_intra_b %>% 
  pull(var) %>% 
  mean()
tegula_n <- quad_intra_b %>% 
  pull(tot) %>% 
  mean() %>% 
  ceiling() #round up to whole number for simulation
```

Use simulation to run power analysis for comparing tegula across years sites
```{r}
n.sim <- 10000         ## number of simulations
mu    <- tegula_avg  ## site means
sigma2<- tegula_var  ## error variance (averaged across all sites) assumption: 7.5
n     <- tegula_n    ## number of observations per site (mean across all years: 9)

n     <- 24           # (if I am running two sets of surveys in a year)
g     <- length(mu) 
group <- factor(rep(LETTERS[1:g], each = n))

results <- numeric(n.sim) ## vector to store results in

for(i in 1:n.sim){
  ## Simulate new response, build data set
  y <- rnorm(n * g, mean = rep(mu, each = n), sd = sqrt(sigma2))
  data <- data.frame(y = y, group = group)
 
  ## Fit one-way ANOVA model
  fit  <- aov(y ~ group, data = data)
  
  ## Extract result of global F-test
  results[i] <- summary(fit)[[1]][1, "Pr(>F)"] < 0.05 ## 1 = reject
}

mean(results) ## proportion of simulation runs that rejected H_0
```
n = 24 for 80% power at BML

## Alegria 
```{r}
quad_intra_a <- quad %>%
  filter(species %in% c("tegula funebralis"),
         site == "Alegria") %>% 
  group_by(year) %>% 
  summarize(tot=n(),
            avg=mean(count),
            var=var(count))

tegula_avg <- quad_intra_a %>% 
  pull(avg)
tegula_var <- quad_intra_a %>% 
  pull(var) %>% 
  mean()
tegula_n <- quad_intra_a %>% 
  pull(tot) %>% 
  mean() %>% 
  ceiling() #round up to whole number for simulation
```

Use simulation to run power analysis for comparing tegula across years
```{r}
n.sim <- 10000         ## number of simulations
mu    <- tegula_avg  ## site means
sigma2<- tegula_var  ## error variance (averaged across all sites) assumption: 7.5
n     <- tegula_n    ## number of observations per site (mean across all years: 15)

n     <- 28           # (if I am running two sets of surveys in a year)
g     <- length(mu) 
group <- factor(rep(LETTERS[1:g], each = n))

results <- numeric(n.sim) ## vector to store results in

for(i in 1:n.sim){
  ## Simulate new response, build data set
  y <- rnorm(n * g, mean = rep(mu, each = n), sd = sqrt(sigma2))
  data <- data.frame(y = y, group = group)
 
  ## Fit one-way ANOVA model
  fit  <- aov(y ~ group, data = data)
  
  ## Extract result of global F-test
  results[i] <- summary(fit)[[1]][1, "Pr(>F)"] < 0.05 ## 1 = reject
}

mean(results) ## proportion of simulation runs that rejected H_0
```
n = 28 for 80% power at Alegria

## Cabrillo 
```{r}
quad_intra_c <- quad %>%
  filter(species %in% c("tegula funebralis"),
         site == "Cabrillo III") %>% 
  group_by(year) %>% 
  summarize(tot=n(),
            avg=mean(count),
            var=var(count))

tegula_avg <- quad_intra_c %>% 
  pull(avg)
tegula_var <- quad_intra_c %>% 
  pull(var) %>% 
  mean()
tegula_n <- quad_intra_c %>% 
  pull(tot) %>% 
  mean() %>% 
  ceiling() #round up to whole number for simulation
```

Use simulation to run power analysis for comparing tegula across years
```{r}
n.sim <- 10000         ## number of simulations
mu    <- tegula_avg  ## site means
sigma2<- tegula_var  ## error variance (averaged across all sites) assumption: 7.5
n     <- tegula_n    ## number of observations per site (mean across all years: 6)

n     <- 9           # (if I am running two sets of surveys in a year)
g     <- length(mu) 
group <- factor(rep(LETTERS[1:g], each = n))

results <- numeric(n.sim) ## vector to store results in

for(i in 1:n.sim){
  ## Simulate new response, build data set
  y <- rnorm(n * g, mean = rep(mu, each = n), sd = sqrt(sigma2))
  data <- data.frame(y = y, group = group)
 
  ## Fit one-way ANOVA model
  fit  <- aov(y ~ group, data = data)
  
  ## Extract result of global F-test
  results[i] <- summary(fit)[[1]][1, "Pr(>F)"] < 0.05 ## 1 = reject
}

mean(results) ## proportion of simulation runs that rejected H_0
```
n = 9 for 80% power at Cabrillo III

# Run power analysis on Tegula at each year within Alegria

## 2001
```{r}
quad_intra_2001 <- quad %>%
  filter(species %in% c("tegula funebralis"),
         site == "Alegria",
         year == "2001") %>% 
  group_by(transect) %>% 
  summarize(tot=n(),
            avg=mean(count),
            var=var(count)) %>%
  add_row(transect = 3, tot = 0, avg = 0, var = 0) %>% 
  replace(is.na(.), 0)
  

tegula_avg <- quad_intra_2001 %>% 
  pull(avg)
tegula_var <- quad_intra_2001 %>% 
  pull(var) %>% 
  mean()
tegula_n <- quad_intra_2001 %>% 
  pull(tot) %>% 
  mean() %>% 
  ceiling() #round up to whole number for simulation
```

## 2012
```{r}
quad_intra_2012 <- quad %>%
  filter(species %in% c("tegula funebralis"),
         site == "Alegria",
         year == "2012") %>% 
  group_by(transect) %>% 
  summarize(tot=n(),
            avg=mean(count),
            var=var(count)) %>%
  replace(is.na(.), 0)

tegula_avg <- quad_intra_2012 %>% 
  pull(avg)
tegula_var <- quad_intra_2012 %>% 
  pull(var) %>% 
  mean()
tegula_n <- quad_intra_2012 %>% 
  pull(tot) %>% 
  mean() %>% 
  ceiling() #round up to whole number for simulation
```

Use simulation to run power analysis for comparing tegula across years
```{r}
n.sim <- 10000         ## number of simulations
mu    <- tegula_avg  ## transect means
sigma2<- tegula_var  ## error variance (averaged across all transects) assumption: 7.5
n     <- tegula_n    ## number of observations per transect (n=3 for 2012, n=2 for 2001)

n     <- 2           # (if I am running two sets of surveys in a year)
g     <- length(mu) 
group <- factor(rep(LETTERS[1:g], each = n))

results <- numeric(n.sim) ## vector to store results in

for(i in 1:n.sim){
  ## Simulate new response, build data set
  y <- rnorm(n * g, mean = rep(mu, each = n), sd = sqrt(sigma2))
  data <- data.frame(y = y, group = group)
 
  ## Fit one-way ANOVA model
  fit  <- aov(y ~ group, data = data)
  
  ## Extract result of global F-test
  results[i] <- summary(fit)[[1]][1, "Pr(>F)"] < 0.05 ## 1 = reject
}

mean(results) ## proportion of simulation runs that rejected H_0
```
n = 7 transects for 2012 to capture significant variation at Alegria ; n = 2 (?!) for 2001


