---
title: "Summer survey - swath/quadrat combination"
author: "Amelia Ritger"
date: "8/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# RUN THIS AFTER RUNNING "summer-swath.Rmd", "summer-quadrat.Rmd"

Compare densities of Pachygraspus counted from swaths vs quadrats for each site
```{r}
quadrat_pachy <- quadrat_sum %>%
  filter(species=="Small crabs") %>%
  mutate(species=str_replace(species, "Small crabs", "quadrats"))

swath_pachy <- swath_sum %>%
  filter(species=="Small crabs") %>%
  mutate(species=str_replace(species, "Small crabs", "swaths"))

pachy_comp <- full_join(quadrat_pachy, swath_pachy) %>%
  rename(type=species)

#set site order for plotting
pachy_comp$location <- factor(pachy_comp$location, levels=c("Cabrillo NM", "Alegria", "Lompoc Landing", "Bodega Bay", "Fogarty Creek", "Sokol Point"))

ggplot(pachy_comp, aes(x=location, y=density_per_m2, fill=type)) +
  geom_bar(position="dodge", stat="identity")

ggsave(here("figures", "pachygraspus-comparison.png"), height=20, width=40, units="cm")
```

OK, so let's pretend that we didn't count Pachygraspus on the swaths...

Combine swath and quadrat data into one df (but remove swath pachygraspus counts)
```{r}
quadrat_sum_no_pachy <- quadrat_sum %>%
  filter(!species=="Small crabs") %>%
  mutate(species=str_replace(species, "Whelks", "Small whelks")) %>% #separate kelletia from nucella/acanthinucella
  rename(number_sampled=number_quadrats_sampled) #rename for joining dfs

swath_sum_join <- swath_sum %>%
  rename(number_sampled=number_transects_sampled) #rename for joining dfs

#join the two df
all <- full_join(swath_sum_join, quadrat_sum_no_pachy)

#join "Pisaster" and "Other stars" into "Large stars" and get final summed df
all_sum <- all %>%
  mutate(species=str_replace(species, "Other stars", "Not Pisaster, not small stars"), #set up values for combining "other stars" with "pisaster"
         species=ifelse(str_detect(species, pattern="Pisaster"), "Large stars", species)) %>%
  group_by(location, species) %>%
  summarize(density_per_m2=sum(density_per_m2)) #sum densities together

#set species order for plotting
#all_sum$species <- factor(all_sum$species, levels=c("Tegula funebralis", "Small whelks", "Large whelks", "Small crabs", "Large crabs", "Small stars", "Large stars", "Octopus"))
                                            
#set site order for plotting
all_sum$location <- factor(all_sum$location, levels=c("Cabrillo NM", "Alegria", "Lompoc Landing", "Bodega Bay", "Fogarty Creek", "Sokol Point"))

ggplot(all_sum, aes(x=location, y=density_per_m2, fill=species)) +
  geom_bar(position="dodge", stat="identity")
```

Woof. Let's try a bit of combining

Check out which groups can/should be combined
```{r}
#check crabs first
crabs <- all_sum %>%
  filter(species=="Small crabs" | species=="Large crabs")

ggplot(crabs, aes(x=location, y=density_per_m2, fill=species)) +
  geom_bar(position="dodge", stat="identity")

#Uhh... not a lot of big crabs in general. Looks like I can combine large and small crabs?

#whelks I'm going to guess yes, combine them, since Kelletia were only found at Cabrillo NM

whelks <- all_sum %>%
  filter(species=="Small whelks" | species=="Large whelks")

ggplot(whelks, aes(x=location, y=density_per_m2)) +
  geom_bar(position="dodge", stat="identity") 

#yup. Now onto stars?

stars <- all_sum %>%
  filter(species=="Small stars" | species=="Large stars")

ggplot(stars, aes(x=location, y=density_per_m2, fill=species)) +
  geom_bar(position="dodge", stat="identity") 

#small stars and large stars show interesting trends, so let's keep them separate
```

Combine 'em
```{r}
all_group <- all_sum %>%
  mutate(species=ifelse(str_detect(species, pattern="crabs"), "Crabs", 
                        ifelse(str_detect(species, pattern="whelks"), "Whelks", species))) %>%
  group_by(location, species) %>%
  summarize(density_per_m2=sum(density_per_m2)) %>% #sum densities together
  filter(!species=="Octopus") #remove octopus because they're not doing much :(

#set species order for plotting
all_group$species <- factor(all_group$species, levels=c("Tegula funebralis", "Whelks", "Crabs", "Small stars", "Large stars", "Octopus"))

ggplot(all_group, aes(x=location, y=density_per_m2, fill=species)) +
  geom_bar(position="dodge", stat="identity") +
  ylab(bquote('Species density ' (per ~m^2))) +
  xlab("Location") +
  theme_bw()

ggsave(here("figures", "summer-all.png"), height=20, width=40, units="cm")
```

Plot relative frequencies of predator species for each site
```{r}
ggplot(all_group, aes(x=location, y=density_per_m2, fill=species)) +
  geom_bar(position="fill", stat="identity") + #facet_wrap(~year)
  ylab("Species proportion") +
  xlab("Location") +
  theme_bw()

ggsave(here("figures", "summer-all_proportion.png"), height=20, width=40, units="cm")
```