---
title: "Summer survey - swath/quadrat combination"
author: "Amelia Ritger"
date: "8/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(gt) #For presentable table(s)
```

# RUN THIS AFTER RUNNING "summer-swath.Rmd", "summer-quadrat.Rmd"

Compare densities of Pachygraspus counted from swaths vs quadrats for each site
```{r}
quadrat_pachy <- quadrat_sum %>%
  filter(species=="Small crabs") %>%
  mutate(type=str_replace(species, "Small crabs", "quadrats"))

swath_pachy <- swath_sum %>%
  filter(species=="Small crabs") %>%
  mutate(type=str_replace(species, "Small crabs", "swaths"))

pachy_comp <- full_join(quadrat_pachy, swath_pachy)

#set site order for plotting
pachy_comp$location <- factor(pachy_comp$location, levels=c("Cabrillo NM", "Alegria", "Lompoc Landing", "Bodega Bay", "Fogarty Creek", "Sokol Point"))

ggplot(pachy_comp, aes(x=location, y=density_per_m2, fill=type)) +
  geom_bar(position="dodge", stat="identity") +
  ylab(bquote('Species density ' (per ~m^2))) +
  xlab("Location") +
  theme_bw()  +
  theme(axis.text.x=element_text(size=12),
        axis.title.x=element_text(size=15),
        axis.text.y=element_text(size=12),
        axis.title.y=element_text(size=15),
        legend.text=element_text(size=12),
        legend.title=element_text(size=15))

ggsave(here("figures", "surveys", "pachygraspus-comparison.png"), height=20, width=40, units="cm")
```

OK, so let's pretend that we didn't count Pachygraspus on the swaths...

Combine swath and quadrat data into one df (but remove swath pachygraspus counts)
```{r}
quadrat_sum_no_pachy <- quadrat_sum %>%
  filter(!species=="Small crabs") %>%
  mutate(species=str_replace(species, "Whelks", "Small whelks")) %>% #separate kelletia from nucella/acanthinucella
  dplyr::rename(number_sampled=number_quadrats_sampled) #rename for joining dfs

swath_sum_join <- swath_sum %>%
  dplyr::rename(number_sampled=number_transects_sampled) #rename for joining dfs

#join the two df
all <- full_join(swath_sum_join, quadrat_sum_no_pachy)

#join "Pisaster" and "Other stars" into "Large stars" and get final summed df
all_sum <- all %>%
  mutate(species=str_replace(species, "Other stars", "Not Pisaster, not small stars"), #set up values for combining "other stars" with "pisaster"
         species=ifelse(str_detect(species, pattern="Pisaster"), "Large stars", species)) %>%
  group_by(location, species) %>%
  summarize(density_per_m2=sum(density_per_m2)) #sum densities together

#set species order for plotting
#all_sum$species <- factor(all_sum$species, levels=c("Tegula funebralis", "Small whelks", "Large whelks", "Small crabs", "Large crabs", "Small stars", "Large stars", "Octopus"))
                                            
#set site order for plotting
all_sum$location <- factor(all_sum$location, levels=c("Cabrillo NM", "Alegria", "Lompoc Landing", "Bodega Bay", "Fogarty Creek", "Sokol Point"))

ggplot(all_sum, aes(x=location, y=density_per_m2, fill=species)) +
  geom_bar(position="dodge", stat="identity")
```

Woof. Let's try a bit of combining

Check out which groups can/should be combined
```{r}
#check crabs first
crabs <- all_sum %>%
  filter(species=="Small crabs" | species=="Large crabs")

ggplot(crabs, aes(x=location, y=density_per_m2, fill=species)) +
  geom_bar(position="dodge", stat="identity")

#Uhh... not a lot of big crabs in general. Looks like I can combine large and small crabs?

#whelks I'm going to guess yes, combine them, since Kelletia were only found at Cabrillo NM

whelks <- all_sum %>%
  filter(species=="Small whelks" | species=="Large whelks")

ggplot(whelks, aes(x=location, y=density_per_m2)) +
  geom_bar(position="dodge", stat="identity") 

#yup. Now onto stars?

stars <- all_sum %>%
  filter(species=="Small stars" | species=="Large stars")

ggplot(stars, aes(x=location, y=density_per_m2, fill=species)) +
  geom_bar(position="dodge", stat="identity")

#small stars and large stars show interesting trends, so let's keep them separate
```

Combine 'em
```{r}
all_group <- all_sum %>%
  mutate(species=ifelse(str_detect(species, pattern="crabs"), "Crabs", 
                        ifelse(str_detect(species, pattern="whelks"), "Whelks", species))) %>%
  group_by(location, species) %>%
  summarize(density_per_m2=sum(density_per_m2)) %>% #sum densities together
  filter(!species=="Octopus") #remove octopus because they're not doing much :(

#set species order for plotting
all_group$species <- factor(all_group$species, levels=c("Tegula funebralis", "Whelks", "Crabs", "Small stars", "Large stars", "Octopus"))

ggplot(all_group, aes(x=location, y=density_per_m2, fill=species)) +
  geom_bar(position="dodge", stat="identity") +
  ylab(bquote('Species density ' (per ~m^2))) +
  xlab("Location") +
  theme_bw() +
  theme(axis.text.x=element_text(size=12),
        axis.title.x=element_text(size=15),
        axis.text.y=element_text(size=12),
        axis.title.y=element_text(size=15),
        legend.text=element_text(size=12),
        legend.title=element_text(size=15))

ggsave(here("figures", "surveys", "summer-all.png"), height=20, width=40, units="cm")
```

Plot relative frequencies of predator species for each site
```{r}
ggplot(all_group, aes(x=location, y=density_per_m2, fill=species)) +
  geom_bar(position="fill", stat="identity") + #facet_wrap(~year)
  ylab("Species proportion") +
  xlab("Location") +
  theme_bw() +
  theme(axis.text.x=element_text(size=12),
        axis.title.x=element_text(size=15),
        axis.text.y=element_text(size=12),
        axis.title.y=element_text(size=15),
        legend.text=element_text(size=12),
        legend.title=element_text(size=15))

ggsave(here("figures", "surveys", "summer-all_proportion.png"), height=20, width=40, units="cm")
```


```{r}
swath_prep <- swath %>%
  select(location, date, cancer_sp:octopus)

quad_prep <- quadrat %>%
  select(location, date, tegula_funebralis:leptasterias)

all_raw <- full_join(swath_prep, quad_prep) %>%
  dplyr::rename(pisaster_sp=pisaster,
                henricia_sp=henricia,
                kelletia_kelletii=kelletia,
                mexacanthina_lugubris=mexacanthina,
                acanthinucella_sp=acanthinucella,
                pachygrapsus_crassipes=pachygraspus,
                leptasterias_sp=leptasterias) %>%
  pivot_longer(cols=cancer_sp:leptasterias_sp,
               names_to = "species",
               values_to = "count") %>%
  mutate(species=str_replace(species,"sp", "spp")) %>%
  filter(!species=="first_pisaster_m",
         !species=="cancer_spp",
         count>0,
         date > "2021-6-20") %>% #only look at summer surveys
  select(location, species) %>%
  mutate(location=str_replace_all(location, "Bodega Bay", "Bodega"),
         location=str_replace_all(location, "Bodega", "Bodega Bay")) %>% #I'm tired so I can't figure out a simpler way to do this
  distinct() %>% #get distinct number of location and species combinations
  mutate(species=str_replace(species, "_", " "), #remove underscore from organism names
         species = sub("(.)", "\\U\\1", species, perl=TRUE), #capitalize organism names
         species=str_replace(species, "_", "/"),  #%>% #replace underscore for ostrina/emarginata
         empty="")

#Set site order
all_raw$location <- factor(all_raw$location, levels=c("Cabrillo NM", "Alegria", "Lompoc Landing", "Bodega Bay", "Fogarty Creek", "Sokol Point"))

#Set species order
all_raw$species <- factor(all_raw$species, levels=c("Tegula funebralis", "Nucella ostrina/emarginata", "Nucella canaliculata", "Acanthinucella spp", "Mexacanthina lugubris", "Kelletia kelletii", "Pachygrapsus crassipes", "Cancer antennarius", "Cancer productus", "Cancer oregonensis", "Leptasterias spp", "Henricia spp", "Pisaster spp","Octopus"))

all_raw %>%
  arrange(location, species) %>%
  #dplyr::rename(rowname=location) %>%
  gt(
    groupname_col="location",
    rowname_col="empty") %>%
  cols_align(
    align = "right",
    columns = species) %>%
  cols_label(
    species = "") %>%
  gtsave(here("figures", "surveys", "species_present.png"))

#Create table of all observed species

all_sp <- all_raw %>%
  select(species) %>%
  distinct()

all_sp %>%
  arrange(species) %>%
  #dplyr::rename(rowname=location) %>%
  gt() %>%
  cols_align(
    align = "left",
    columns = species) %>%
  cols_label(
    species = "Observed species") %>%
  gtsave(here("figures", "surveys", "species_all.png"))
```

