---
title: "LOL cycle analysis"
author: "Amelia Ritger"
date: "11/1/2021"
output: html_document
---

Zoom in on LOL as an example of pH variation within tidepool

#Load packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(lubridate)
library(plotly) #create plots with multiple y axes
library(highcharter) #create highcart to interact with data
library(gt) #create gt table
library(scales) #for loess plot axis edits
```

#Load data
```{r}
lol_cycle <- read_csv(here("data","sensors", "lol_all.csv"))
```
 
## Plot up all data from LOL to show extent of pH and temperature drops within pool - code from Li (thank you)
```{r}
figure <- function(df_input,name) {
  
m <- list(
  l = 10,
  r = 10,
  b = 30,
  t = 30,
  pad = 4
)

fig <- plot_ly(width = 1800)
 
 fig <- fig %>% 
   add_trace(data = df_input, y=~p_h, x = ~date_time, name = "pH", yaxis = "y1", mode = "lines", type = "scatter", line=list(color="green")) 

y2 <- list(
  tickfont = list(color = "red", size=20),
  titlefont = list(color = "red", size=30),
  overlaying = "y",
  side = "left",
  anchor="free",
  position=0.02,
  title = "Temperature (\u00B0C)")

fig <- fig %>% 
  add_trace(data =df_input, y=~temp_c, x = ~date_time, name = "Temperature (\u00B0C)", yaxis = "y2", mode = "lines", type = "scatter", line=list(color="red"))

fig <- fig %>% 
  layout(
    title = list(text=name,xanchor = 'center', yanchor =  'top'),
    font= list(size = 20),
    yaxis2 = y2, #yaxis3 = y3,
    xaxis = list(title = 'Date',
                 domain = c(0.1, 1),
                 tick0 = "2021-06-14 08:00:00",
                 dtick = 7*86400000.0, 
                 tickformat="%d-%b",
                 titlefont = list(size = 30),
                 tickfont = list(size = 20)),
    yaxis = list(title = "pH",
              tickfont = list(color = "green", size=20),
              titlefont = list(color = "green", size=30),
              side="left",
              anchor="free",
              position=0.08),
  showlegend = FALSE,
  margin = m)

return(fig)

}

#save it

lol_cycle_fig <- figure(lol_cycle,"Lompoc Landing")

#orca(lol_cycle_fig, file="lol-cycle-full.png") #save plotly image
#filesstrings::move_files("lol-cycle-full.png", here("figures", "sensors"), overwrite=TRUE) #move the image to the "Figures" folder
```
Alternatively, check out https://plotly.com/r/multiple-axes/

## Zoom in on that period of time where pH jumped from 8.2 to 7.2
```{r}
lol_cycle_zoom <- lol_cycle %>%
  filter(date_time < ymd_hms("2021-08-03 00:00:00"),
         date_time > ymd_hms("2021-07-29 00:00:00"))

y1 <- lol_cycle_zoom$temp_c
y2 <- lol_cycle_zoom$temp_c
y3 <- lol_cycle_zoom$tide
x <- lol_cycle_zoom$date_time

highchart() %>% 
  hc_add_series(data = y1, dashStyle="solid") %>% 
  hc_add_series(data = y2, yAxis = 1) %>% 
  #hc_add_series(data = y3, yAxis = 1) %>%
  hc_yAxis_multiples(
     list(lineWidth = 3, lineColor='#D55E00', title=list(text="Temp")),
     #list(lineWidth = 3, lineColor="#009E73", title=list(text="pH")),
     list(lineWidth = 3, lineColor="#0072B2", title=list(text="pH"))) %>%
    hc_xAxis(title = "Date", categories = x, breaks=10) %>%
  hc_colors(c("#D55E00",
              "#009E73",
              "#0072B2"))
```

#Create low pass loess filtering figure
```{r}
lol_prep <- lol_cycle %>%
  filter(tide=="high") %>%
  select(date_time, temp_c, p_h) %>%
  pivot_longer(cols=temp_c:p_h,
               names_to = "group",
               values_to = "value")

count <- 1
breaks_fun <- function(x) {
  count <<- count + 1L
  switch(
    count,
    c(7.2, 7.6, 8.0, 8.6),
    c(5, 10, 15, 20),
  )
}

#Set x axis limits
lims <- as.POSIXct(strptime(c("2021-06-13 00:00:00", "2021-10-09 23:45:00"), 
                   format = "%Y-%m-%d %H:%M:%S"))

ggplot(lol_prep, aes(x=date_time, y=value, group=group)) +
  geom_line(aes(color=group), size=0.7, alpha=0.6) +
  geom_smooth(aes(group=group, color=group), method="loess", span=0.1) + 
  scale_color_manual(values = c("#117733", "#882255")) + #color lines by custom site color palette
  scale_x_datetime(breaks = date_breaks("1 week"),
                   labels = date_format("%b-%d"), #b makes month label June and not 06
                   limits = lims, #set limits
                   expand = c(0, 0)) + 
  facet_grid(group ~ ., #facet wrap to create one panel for pH and one for temp
             scales = "free",
             switch="both",
             labeller = as_labeller(c(temp_c = "Temperature (C)", p_h = "pH"))) + #customize strip labels
  xlab("Date and time") +
  ylab(NULL) + #remove "Value" from Y axis label
  scale_y_continuous(breaks = scales::pretty_breaks(6)) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, vjust = 1, hjust=1, size=22),
        axis.title.x=element_blank(),
        axis.text.y=element_text(size=22),
        legend.position = "none",
        strip.background = element_blank(), #remove strip background from facet_grid
        strip.text.y = element_text(size = 25),
        strip.placement = "outside") + #place the strip outside of the plot
  guides(colour = guide_legend(override.aes = list(size=5))) #make the legend line wider, for easier viewing

ggsave(here("figures", "sensors", "smooth_lol.png"), height=20, width=40, units="cm")
```

## Create a nice table
```{r}
#create df to look at frequency of low pH conditions
lol_7.8 <- lol_cycle %>%
  mutate(threshold = ifelse(p_h<7.8, "below", "above")) %>%
  group_by(threshold) %>%
  summarize(tot_observations = n())

#Get total number of observations - use double square brackets because it's a tibble
lol_n <- lol_7.8[[1,2]] + lol_7.8[[2,2]]

#Get frequency of pH values below 7.8
lol_freq <- lol_7.8[[2,2]]/lol_n
  
# Calculate min, max, average pH and temp values
plol_min <- (min(lol_cycle$p_h))
plol_max <- (max(lol_cycle$p_h))
plol_median <- (median(lol_cycle$p_h))
plol_mean <- (mean(lol_cycle$p_h))
plol_sd <- (sd(lol_cycle$p_h))
plol_cv <- (plol_sd/plol_mean)*100 #coefficient of variation

tlol_min <- (min(lol_cycle$temp_c))
tlol_max <- (max(lol_cycle$temp_c))
tlol_median <- (median(lol_cycle$temp_c))
tlol_mean <- (mean(lol_cycle$temp_c))
tlol_sd <- (sd(lol_cycle$temp_c))
tlol_cv <- (tlol_sd/tlol_mean)*100 #coefficient of variation

#combine into a table
ph_table <- tribble(
  ~"Measurement", ~"Min", ~"Max", ~"Median",  ~"CV", ~"pH <7.8", 
  "pH", plol_min, plol_max, plol_median, plol_cv, lol_freq,
  "Temperature", tlol_min, tlol_max, tlol_median, tlol_cv, NA
  )

#make the table look nice
ph_table %>%
  gt() %>%
  fmt_number(
  columns = c("Min":"Median"),
  rows = everything(),
  decimals = 2) %>%
  data_color(
    columns = c("Measurement"),
    colors = scales::col_factor( # <- bc it's a factor
      palette = c("#117733","#882255"),
      domain = c("pH", "Temperature"))) %>%
  gtsave(here("figures", "sensors", "table_lol.png"))
```

## Next steps:
1. assess the greatest pH and temperature range in a 24 hour, 1 week period and monthly averages for LOL

2. Re-filter all pH data based on HOBO temp data (especially, make annotations next to every single point/range that was filtered out, for future QC)
